Long-range correlations and their event-scale dependence in high-multiplicity pp collisions at $\sqrt{s} = 13$~TeV

### 2020-10-24 sec_results.tex
updates

Figure~\ref{fig:PlotCorrMBHMT} shows two-particle correlation functions as a function of the $\Delta \eta$ and $\Delta \varphi$ separation of trigger and associated particles. The per trigger yield iss measured with Eq.~\ref{eq:corrfunction} for $1 < \pttrig~\mathrm (\ptassoc) < 2$ GeV/$c$ in pp collisions at $\sqrt{\it{s}} = $\unit{13} {\rm{}TeV}. The left panel is for the 0--0.1\% multiplicity class. The right one is from the minimum bias events (0--100\% multiplicity class). The ridge structure is clearly observed in the high-multiplicity class while it is not seen in the minimum bias events. The away-side yield is populated mostly by correlation of jets which are back-to-back.

@@

Projected $\Delta\varphi$ distributions of the two-particle correlation functions in $1.6<|\Delta\eta|<1.8$ are shown in Fig.~\ref{fig:PlotDeltaPhi} for various $\it{p}_{\rm{T}}$ intervals in the minimum bias class (0--100\%) and the high-multiplicity class (0--0.1\%). The near-side ($\Delta\varphi\sim 0$) peak in 0--0.1\% multiplicity class is clearly observed in all transverse momentum ranges studied while there is no hint of the signal in the minimum bias class. The near-side peak is highest in the interval of $1<\it{p}_{\rm{T}}<2$ GeV/$c$ and gradually decreases with increasing $\it{p}_{\rm{T}}$ in the high-multiplicity class. The measurements in the high-multiplicity class are compared to the results measured by the CMS collaboration~\cite{Khachatryan:2015lva}. The near-side peaks in all transverse momentum ranges are comparable to each other within the uncertainties. The larger away yields observed in \cite{Khachatryan:2015lva} can be attributed to the difference in $\eta$ acceptances. The model predictions are presented for the measurements in the high-multiplicity class with the same acceptances used in this article both for $\eta$ and event multiplicity. The high-multiplicity class for each model is determined by requiring the minimum number of charged particles larger than 105, which corresponds to the top 0--0.1\% multiplicity class based on $\pythiam$. The $\pythiashoving$ gives good estimates of the near-side peak and slightly overestimates the away-side yield for the interval of $1<\it{p}_{\rm{T}}<2$ GeV/$c$. However the $\pythiashoving$ underestimates the near-side peak for the $\it{p}_{\rm{T}}>2$ GeV/$c$ range. $\pythiam$ does not show any peaks in the near-side as expected. It slightly underestimates the away-side peak for $1<\it{p}_{\rm{T}}<2$ GeV/$c$ and gives good estimates for the $\it{p}_{\rm{T}}>2$ GeV/$c$ range. On the other hand, $\epos$ describes $\pt$ dependence quantitatively for the $1<\it{p}_{\rm{T}}<4$ GeV/$c$ range., while overestimating the near-side peak for the $\it{p}_{\rm{T}}>2$ GeV/$c$ range.

\begin{figure}[h!]
	\centering
	\subfigure{\includegraphics[width=0.99\textwidth]{./figures/Fig3_PlotRidgeYield.eps}}
	\caption{ The spectra of ridge yield as function of transverse momentum. The filled circles denote the measurement with ALICE and compared to the CMS measurement~\cite{Khachatryan:2015lva}, which is represented as open blue box. The three lines show model predictions from $\pythiam$ (blue line), $\pythiashoving$ (orange line) and $\epos$ (green line).}
	\label{fig:PlotYSpect}
\end{figure}

Fig.~\ref{fig:PlotYSpect} shows the near-side ridge yield measured in high-multiplicity events as a function of transverse momentum. The ALICE data are compared to \cite{Khachatryan:2015lva}.
%The ridge yields as a function of the transverse momentum are shown in Fig.~\ref{fig:PlotYSpect} in the high-multiplicity class and compared to \cite{Khachatryan:2015lva}.
The particle multiplicity in our analysis is estimated with the forward V0 subsystem, whereas whereas the analysis \cite{Khachatryan:2015lva}  used the mid-rapidity particles in $|\eta|<$2.4 and $\pt>0.4$ GeV/$c$. By interpolating the $\eta$ and $\pt{}$ distributions based on PYTHIA, it is found that the multiplicity in \cite{Khachatryan:2015lva} is about 20\% higher than 0--0.1\% in this analysis, corresponding 0--0.03\%. Taking into account the differences in acceptance of charged tracks and estimated multiplicity, the measurements are comparable with each other. The spectrum is also compared with model calculations. $\pythiam$ gives zero yields since it does not contain the ridge effect. $\pythiashoving$ describes the yield qualitatively however its yield decreases more rapidly than the measured data as $\pttrigassoc$ increases. $\epos$, unlike $\pythiashoving$, gives estimates for the $\pt$ dependence showing the ridge yield for the $\it{p}_{\rm{T}}>2$ GeV/$c$ range, while estimating larger yield for the $\it{p}_{\rm{T}}<2$ GeV/$c$.

@@

The ridge yield is further studied with respect to two different event-scales of the hard processes. Event-scale measurement by the largest $\pt$ of charged particles in each event is denoted as $\ptlead$ while the other is done by reconstructing jets and measuring their transverse momenta, represented as $\ptjet$. As shown in Fig.~\ref{fig:PlotCorrHMTSel}, the ridge structure for $1< \pttrig\,(\ptassoc) <2$ GeV/$c$ still persists in high-multiplicity pp collisions with $\ptlead>9$ (left) or $\ptjet>10$~GeV/$c$ (right) .  %The interval is the same to the measured ridge yield is maximum without the event-scale selection.
It is worth noting that the correlation function with the minimum $\ptjet$ selection has double peaks along the $\Delta\eta$ axis at $\Delta\varphi = \pi$ due to the limited jet acceptance by requiring a minimum $\ptjet$.

@@

Figure~\ref{fig:PlotDeltaPhiESE} shows projected $\Delta\varphi$ distributions of the correlation functions in $1.6<|\Delta\eta|<1.8$ for the minimum $\ptlead$ (lower) and $\ptjet$ (upper) requirement. Even with the event-scale selection, the ridge is still visible at the near-side. The near-side yield increases as the threshold of $\ptjet$ increases compared to one measured in unbiased events, while the ridge yields for an increasing threshold of $\ptlead$ increase less than those of $\ptjet$. The event-scale measurement with $\ptlead$ favors harder jets as a leading particle carries a large fraction of the momentum. In contrast, soft jets are favored for reconstructed jets. The results are compared with $\pythiashoving$ and $\pythiam$ calculations. $\pythiashoving$ can still reproduce the near-side peak, while underestimating the yield with the event-scale selections. On the other hand, $\pythiam$ does not show the near-side peak with the event-scale selections. Both models describe the away-side yield with the $\ptjet$ selection and overestimate away-side yield with the $\ptlead$ selection.

@@

Finally, the ridge yields as functions of the minimum $\ptlead$ ($\it{p}^{\rm{LP}}_{\rm{T,min}}$) and $\ptjet$ ($\it{p}^{\rm{Jet}}_{\rm{T,min}}$) selections are shown in Fig.~\ref{fig:RidgeYield_ESE}. The ridge yields are increased for $\it{p}^{\rm{LP}}_{\rm{T,min}} >$ 7 and $\it{p}^{\rm{Jet}}_{\rm{T,min}} \ge$ 10~GeV/$c$ compared to one from the unbiased events. A moderate increase of the ridge yields toward higher $\ptlead$ or $\ptjet$ is observed and there is no difference between two event selections within the uncertainties.
The results are compared with $\pythiashoving$ and $\pythiam$. $\pythiashoving$ doesn't show a clear $\it{p}^{\rm{LP}}_{\rm{T,min}}$ or $\it{p}^{\rm{Jet}}_{\rm{T,min}}$ dependence within the errors and underestimates the yields especially for the jet tagged events. $\pythiam$ does not produce the ridge yield, estimating zero yield, which implies negligible jet contamination in the measurements.
The origin of the enhanced ridge yields for higher momentum jet-tagged events is not clear to date but it might be related with the expected shorter impact parameters for dijet or multi-jet production events studied in~\cite{Frankfurt:2010ea}.

>>>>

Figure~\ref{fig:PlotCorrMBHMT} shows two-particle correlation functions as a function of the $\Delta \eta$ and $\Delta \varphi$ separation of trigger and associated particles. The per trigger yield iss measured with Eq.~\ref{eq:corrfunction} for $1 < \pttrig~\mathrm (\ptassoc) < 2$ GeV/$c$ in pp collisions at $\sqrt{\it{s}} = $\unit{13} {\rm{}TeV}. The left panel is for the high-multiplicity class. The right one is from the minimum bias events. The ridge structure is clearly observed in the high-multiplicity class while it is not seen in the minimum bias events. The away-side yield is populated mostly by correlation of jets which are back-to-back.

@@

Projected $\Delta\varphi$ distributions of the two-particle correlation functions in $1.6<|\Delta\eta|<1.8$ are shown in Fig.~\ref{fig:PlotDeltaPhi} for various $\it{p}_{\rm{T}}$ intervals in the minimum bias class and the high-multiplicity class. The near-side ($\Delta\varphi\sim 0$) peak in high-multiplicity class is clearly observed in all transverse momentum ranges studied while there is no hint of the signal in the minimum bias class. The near-side peak is highest in the interval of $1<\it{p}_{\rm{T}}<2$ GeV/$c$ and gradually decreases with increasing $\it{p}_{\rm{T}}$ in the high-multiplicity class. The measurements in the high-multiplicity class are compared to the results measured by the CMS collaboration~\cite{Khachatryan:2015lva}. The particle multiplicity in the present analysis is estimated with the forward V0 subsystem, whereas whereas the analysis \cite{Khachatryan:2015lva}  used the mid-rapidity particles in $|\eta|<$2.4 and $\pt>0.4$ GeV/$c$. By interpolating the $\eta$ and $\pt{}$ distributions based on PYTHIA, it is found that the multiplicity in \cite{Khachatryan:2015lva} is about 20\% higher than 0--0.1\% in this analysis, corresponding 0--0.03\%. The near-side peaks in all transverse momentum ranges are comparable to each other within the uncertainties. The larger away yields observed in \cite{Khachatryan:2015lva} can be attributed to the difference in $\eta$ acceptances. The model predictions are presented for the measurements in high-multiplicity class with the identical charged track acceptance and the $\Delta\eta$ projection range to those used in this article. The high-multiplicity class for each model is determined by requiring the minimum number of charged particles larger than 105, which corresponds to the top 0--0.1\% multiplicity class based on $\pythiam$. The $\pythiashoving$ gives good estimates of the near-side peak and slightly overestimates the away-side yield for the interval of 1$<\it{p}_{\rm{T}}<$2 GeV/$c$. However the $\pythiashoving$ underestimates the near-side peak for the $\it{p}_{\rm{T}}>$2 GeV/$c$ range. $\pythiam$ does not show any peaks in the near-side as expected. It slightly underestimates the away-side peak for $1<\it{p}_{\rm{T}}<$2 GeV/$c$ and gives good estimates for the $\it{p}_{\rm{T}}>$2 GeV/$c$ range. On the other hand, $\epos$ describes better $\pt$ dependence quantitatively for the 1$<\it{p}_{\rm{T}}<$4 GeV/$c$ range, while overestimating the near-side peak for the $\it{p}_{\rm{T}}<$2 GeV/$c$ range.

\begin{figure}[h!]
	\centering
	\subfigure{\includegraphics[width=0.99\textwidth]{./figures/Fig3_PlotRidgeYield.eps}}
	\caption{ Fully corrected near-side ridge yield as a function transverse momentum. The filled circles denote the measurement with ALICE. The CMS measurement~\cite{Khachatryan:2015lva} is represented by open blue box markers. The three lines show model predictions from $\pythiam$ (blue line), $\pythiashoving$ (orange line) and $\epos$ (green line).}
	\label{fig:PlotYSpect}
\end{figure}

Fig.~\ref{fig:PlotYSpect} shows the near-side ridge yield measured in high-multiplicity events as a function of transverse momentum. The ALICE data are compared to \cite{Khachatryan:2015lva}.
%The ridge yields as a function of the transverse momentum are shown in Fig.~\ref{fig:PlotYSpect} in the high-multiplicity class and compared to \cite{Khachatryan:2015lva}.
Taking into account the differences in acceptance of charged tracks and estimated multiplicity, the measurements are comparable with each other. The spectrum is also compared with model calculations. $\pythiam$ gives zero yields since it does not contain the ridge effect. $\pythiashoving$ describes the yield qualitatively however its yield decreases more rapidly than the measured data as $\pttrigassoc$ increases. $\epos$, unlike $\pythiashoving$, gives estimates for the $\pt$ dependence showing the ridge yield for the $\it{p}_{\rm{T}}>2$ GeV/$c$ range, while estimating larger yield for the $\it{p}_{\rm{T}}<2$ GeV/$c$.

@@

The ridge yield is further studied with respect to two different event-scales of the hard processes. Event-scale measurement by the largest $\pt$ of charged particles in each event is denoted as $\ptlead$ while the other is done by reconstructing jets and measuring their transverse momenta, represented as $\ptjet$. As shown in Fig.~\ref{fig:PlotCorrHMTSel}, the ridge structure for $1< \pttrig\,(\ptassoc) <2$ GeV/$c$ still persists in high-multiplicity pp collisions with $\ptlead>9$~GeV/$c$ (left) or $\ptjet>10$~GeV/$c$ (right) .  %The interval is the same to the measured ridge yield is maximum without the event-scale selection.
It is worth noting that the correlation function with the minimum $\ptjet$ selection has double peaks along the $\Delta\eta$ axis at $\Delta\varphi = \pi$ due to the jet tagging with the limited jet acceptance in the mid-rapidity as described in this article.

@@

Figure~\ref{fig:PlotDeltaPhiESE} shows projected $\Delta\varphi$ distributions of the correlation functions in $1.6<|\Delta\eta|<1.8$ with the minimum $\ptlead$ (lower) and $\ptjet$ (upper) requirement. Even with the event-scale selection, the ridge is still visible at the near-side. The near-side yield increases as the threshold of $\ptjet$ increases compared to one measured in unbiased events, while the ridge yields for an increasing threshold of $\ptlead$ increase less than those of $\ptjet$. \iffalse The event-scale measurement with $\ptlead$ favors harder jets as a leading particle carries a large fraction of the momentum. In contrast, soft jets are favored for reconstructed jets. \fi The results are compared with $\pythiashoving$, $\pythiam$, and $\epos$ calculations. The near-side peaks are qualitatively reproduced by $\pythiashoving$ and $\epos$, while underestimated by $\pythiashoving$ and overestimated by $\epos$. On the other hand, $\pythiam$ does not show the near-side peak with the event-scale selections, while giving similar estimates for away-side yield as $\pythiashoving$.
%Both models describe the away-side yield with the $\ptjet$ selection and overestimate away-side yield with the $\ptlead$ selection.

@@

Finally, the ridge yields as functions of the minimum $\ptlead$ ($\it{p}^{\rm{LP}}_{\rm{T,min}}$) and $\ptjet$ ($\it{p}^{\rm{Jet}}_{\rm{T,min}}$) selections are shown in Fig.~\ref{fig:RidgeYield_ESE}. The ridge yields are increased for $\it{p}^{\rm{LP}}_{\rm{T,min}} >$ 7 and $\it{p}^{\rm{Jet}}_{\rm{T,min}} \ge$ 10~GeV/$c$ compared to one from the unbiased events. A moderate increase of the ridge yields toward higher $\ptlead$ or $\ptjet$ is observed and there is no difference between two event selections within the uncertainties.
The results are compared with $\pythiashoving$, $\pythiam$ and $\epos$. $\pythiashoving$ describes comparable trend with data, underestimating the ridge yield. On the other hand, $\epos$ also describes comparable trend with data, overestimating the ridge yield. The origin of the enhanced ridge yields for higher momentum jet-tagged events is not clear to date but it might be related with the expected shorter impact parameters for dijet or multi-jet production events studied in~\cite{Frankfurt:2010ea}.

### 2020-10-24 sec_summary.tex
updates

\section{Conclusions}
\label{sec:summary}
Long-range azimuthal correlations are studied in high-multiplicity pp collisions at $\sqrt{s} = 13$~TeV. These results are obtained in the relative pseudorapidity range $1.6 < |\Delta\eta| < 1.8$ and transverse momentum interval $1 < \pt < 5$~GeV/$c$ for the trigger and associated particles. 
The results are compatible to the measurements by the CMS collaboration~\cite{Khachatryan:2015lva}.
%In addition, the $\pt$ and the $\it{p}_{\rm{T,Lead}}$ or $\it{p}_{\rm{T,Jet}}$ selection dependence of the long-range azimuthal correlations are measured.
$\pythiashoving$ describes the yield qualitatively but its yield is decreased more rapidly than the data as $\pttrigassoc$ increases. On the other hand, $\pythiam$ does not describe the ridge in the long-range as expected. The results from $\pythiam$ also assure that the measured ridge yield is not contaminated by non-flow effects.
Furthermore the ridge yields are studied in events, where jets with a harder fragmentation are preferred by requiring the minimum value of the transverse momentum of leading particles in each event $\ptlead$ or the one of reconstructed jets $\ptjet$. The ridge structure still persist with both selections. The ridge yields increase as $\ptlead$ and $\ptjet$ increase. The latter, however, is more significant than the former selection. The results are compared with $\pythiashoving$ calculations, showing the enhancement while underestimating the absolute ridge yield.
The results might open new way of studying impact parameter dependence of the small systems with jet tagged events in the future and will help to explore possible physical origins of long-range correlations.

>>>>

\section{Conclusions}
\label{sec:summary}
Long-range azimuthal correlations are studied in high-multiplicity pp collisions at $\sqrt{s} = 13$~TeV. These results are obtained in the relative pseudorapidity range $1.6 < |\Delta\eta| < 1.8$ and transverse momentum interval $1 < \pt < 5$~GeV/$c$ for the trigger and associated particles. 
The results are compatible to the measurements by the CMS collaboration~\cite{Khachatryan:2015lva}.
%In addition, the $\pt$ and the $\it{p}_{\rm{T,Lead}}$ or $\it{p}_{\rm{T,Jet}}$ selection dependence of the long-range azimuthal correlations are measured.
$\pythiashoving$ describes the yield qualitatively but its yield is decreased more rapidly than the data as $\pttrigassoc$ increases. On the other hand, $\epos$ gives better description for $\pt$ dependence while overestimating the ridge yield in the low $\pt<$2 GeV/$c$. $\pythiam$ does not describe the ridge in the long-range, which gives closure of the results against non-flow contamination.
%Furthermore the ridge yields are studied in events, where jets with a harder fragmentation are preferred by requiring the minimum value of the transverse momentum of leading particles in each event $\ptlead$ or the one of reconstructed jets $\ptjet$. The ridge structure still persist with both selections. The ridge yields increase as $\ptlead$ and $\ptjet$ increase. The latter, however, is more significant than the former selection. The results are compared with $\pythiashoving$ calculations, showing the enhancement while underestimating the absolute ridge yield.
Furthermore the ridge yields are studied with the event-scale selections through the minimum transverse momentum of jet or leading particle requirement, which are more prone of requiring harder processes in pp collisions. 
The ridge structure still persists with both selections. The ridge yields increase as $\ptlead$ and $\ptjet$ increase. The latter, however, is more significant than the former selection. The results are compared with $\pythiashoving$, $\epos$ and $\pythiam$ calculations, where $\pythiam$ also provides closure against non-flow contamination by estimate ridge yield to be zero with event-scale selections. $\pythiashoving$ and $\epos$ estimate qualitative trends for event-scale selection. The former underestimates the ridge yield, unlike the latter overestimating the ridge yield.
The results might open new way of studying impact parameter dependence of the small systems with jet tagged events in the future and will help to explore possible physical origins of long-range correlations.
